\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\usepackage{wasysym}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{trfsigns}
\usepackage{cmll}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{lipsum}
\usepackage{extpfeil}
\usepackage{soul}
\usepackage{graphicx}

\newcommand\mapsfrom{\mathrel{\reflectbox{\ensuremath{\mapsto}}}}

\theoremstyle{defin}
\newtheorem{defin}{Definition}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{prop}
\newtheorem{prop}{Proposition}

\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}

\theoremstyle{ex}
\newtheorem{ex}{Example}


\theoremstyle{col}
\newtheorem{col}{Corollary}

\author{Daniel Rogozin}
\date{}
\title{Notes on filtration of logics containing {\bf K5}}

\begin{document}

\maketitle

\nocite{*}

\section{Preliminaries}

\begin{defin}
  An $n$-normal modal logic is a set of formulas that contains all Boolean tautologies, formulas $\Diamond_i p \lor \Diamond_i q \leftrightarrow \Diamond_i (p \lor q)$ and $\Diamond_i \bot \leftrightarrow \bot$ for $i \leq n$, and is closed under modus ponens, substitution, and monotonicity: from $\varphi \rightarrow \psi$ infer $\Diamond_i \varphi \rightarrow \Diamond_i \psi$ for $i \leq n$.
\end{defin}

\begin{defin} An $n$-Kripke model is a triple $\mathcal{M} = \langle W, R_1, \dots, R_n, \vartheta \rangle$, where $R_i \subseteq W \times W$, $\vartheta : \operatorname{PV} \to 2^W$, and the connectives have the following semantics:

  \begin{enumerate}
    \item $\mathcal{M}, w \models p \Leftrightarrow w \in \vartheta(p)$
    \item $\mathcal{M}, w \models \not \varphi \Leftrightarrow \mathcal{M}, w \nvDash \varphi$
    \item $\mathcal{M}, w \models \varphi \lor \psi \Leftrightarrow \mathcal{M}, w \models \varphi \text{ or } \mathcal{M}, w \models \psi$
    \item $\mathcal{M}, w \models \Diamond_i \varphi \Leftrightarrow \exists v \in R_i(w) \: \mathcal{M}, v \models \varphi$
  \end{enumerate}
\end{defin}

By ${\bf K5}$ we mean the logic ${\bf K} \oplus {A5}$, where ${A5} = \Diamond p \to \Box \Diamond p$. It is known that ${\bf K5}$ is the modal logic of all Euclidean frames. A frame is called Euclidean if for each $x, y, z$, $x R y$ and $x R z$ implies $y R z$.

\begin{prop}
  Let $\mathcal{F} = \langle W, R \rangle$ be an Euclidean frame.

  \begin{enumerate}
    \item For each $x, y, z \in W$, $x R y$ and $x R z$ implies either $y R z$ or $z R y$.
    \item For each $x \in W$, $R^{*}(x) = \{ x\} \cup R(R(x))$.
    \item $R^{-1} \circ R \subseteq R$.
  \end{enumerate}
\end{prop}

\begin{prop} \label{K5proves}
  ${\bf K5}$ proves
  \begin{enumerate}
    \item $\Box^3 p \leftrightarrow \Box^2 p$
    \item $\Box^2 \Diamond p \leftrightarrow \Box \Diamond p$
    \item $\Box \Diamond \Box p \leftrightarrow \Box \Box p$
    \item $\Box \Diamond^2 p \leftrightarrow \Box \Diamond p$
  \end{enumerate}
\end{prop}

\begin{prop} \label{k5useful}
  Let $\mathcal{M}$ be a ${\bf K}5$ model, $x R y$ for $x, y \in W$ then one has
  \begin{center}
    $\mathcal{M}, x \models \Diamond \Box \varphi$ iff $\mathcal{M}, y \models \Diamond \Box \varphi$.
  \end{center}
\end{prop}

\begin{proof}
$ $

  \begin{enumerate}
    \item Suppose $\mathcal{M}, x \models \Diamond \Box \varphi$. One also has $\mathcal{M}, x \models \Diamond \Box \varphi \to \Box \Diamond \Box \varphi$, so $\mathcal{M}, x \models \Box \Diamond \Box \varphi$. Thus, $\mathcal{M}, y \models \Diamond \Box \varphi$ since $y \in R(x)$.
    \item Suppose $\mathcal{M}, y \models \Diamond \Box \varphi$, then $\mathcal{M}, y \models \Box \varphi$, so $\mathcal{M}, x \models \Diamond \Box \varphi$.
  \end{enumerate}
\end{proof}

\subsection{Filtrations: general definitions}

Let $\mathcal{M} = \langle W, R_1, \dots, R_n, \vartheta \rangle$ be a Kripke model and $\Gamma$ a set of formulas closed under subformulas. An equivalence relation $\sim$ is set to have a finite index if the quotient set $W / \sim$ is finite. The equivalence relation $\sim_{\Gamma}$ induced by $\Gamma$ is defined as

\begin{center}
  $w \sim_{\Gamma} v \Leftrightarrow \forall \varphi \in \Gamma \: (\mathcal{M}, w \models \varphi \Leftrightarrow \mathcal{M}, v \models \varphi)$.
\end{center}

If $\Gamma$ is finite, then $\sim_{\Gamma}$ has a finite index. An equivalence relation $\sim$ respects $\sim_{\Gamma}$, if $w \sim v$ implies $w \sim_{\Gamma} v$.

\begin{defin}
  Let $\mathcal{M} = \langle W, R_1, \dots, R_n, \vartheta \rangle$ be a Kripke model and $\Gamma$ be a Sub-closed set formulas. A $\Gamma$-filtration of $\mathcal{M}$ is a model
  $\widehat{\mathcal{M}} = \langle \widehat{W}, \widehat{R_1}, \dots, \widehat{R_n}, \widehat{\vartheta} \rangle$ such that:
  \begin{enumerate}
    \item $\widehat{W} = W / \sim$, where $\sim$ is an equivalence relation having a finite index that respects $\Gamma$
    \item $\widehat{\vartheta}(p) = \{ [x]_{\sim} \: | \: x \in W \: \& \: x \in \vartheta(p)\}$
    \item For each $i \in I$ one has $\widehat{R}_i^{\text{min}} \subseteq \widehat{R}_i \subseteq \widehat{R}_i^{\text{max}}$. $\widehat{R}_{i, \sim}^{\text{min}}$ is the $i$-th minimal filtered relation on $\widehat{W}$ defined as
    \begin{center}
      $\hat{x} \widehat{R}_{i, \sim}^{\text{min}} \hat{y} \Leftrightarrow \exists x' \sim x \: \exists y' \sim y \: x R_i y$
    \end{center}
    $\widehat{R}_{\Gamma,i}^{\text{max}}$ is the $i$-th maximal filtered relation on $\widehat{W}$ induced by $\Gamma$ defined as
    \begin{center}
      $\hat{x} \widehat{R}_{\Gamma,i}^{\text{max}} \hat{y} \Leftrightarrow \forall \Box_i \varphi \in \Gamma \: (\mathcal{M}, x \models \Box_i \varphi \Rightarrow \mathcal{M}, y \models \varphi)$
    \end{center}
  \end{enumerate}
\end{defin}

If $\Phi$ is finite subset of $\Gamma$ and $\sim = \sim_{\Phi}$, then $\widehat{\mathcal{M}}$ is a definable $\Gamma$-filtration of $\mathcal{M}$ through $\Phi$. If $\sim = \sim_{\Gamma}$, then such a filtration by means of the definiton above is called \emph{strict}.

\begin{lemma}
  Let $\Gamma$ be a finite set of formulas closed under subformulas and $\widehat{\mathcal{M}}$ a filtration of $\mathcal{M}$ through $\Gamma$, then for each $x \in W$ and for each $\varphi \in \Gamma$ one has
  \begin{center}
    $\mathcal{M}, x \models \varphi \Leftrightarrow \widehat{\mathcal{M}}, \hat{x} \models \varphi$
  \end{center}
\end{lemma}

\begin{defin} Let $\mathbb{F}$ be a class of Kripke frames and $\Gamma$ a finite set of formulas closed under subformulas. If for every model $\mathcal{M}$ over $\mathcal{F} \in \mathbb{F}$ there exists a model that is a $\Gamma$-definable filtration of $\mathcal{M}$, then $\mathbb{F}$ admits definable filtration. A class of models $\mathbb{M}$ admits definable filtration if for every $\mathcal{M} \in \mathbb{M}$ there exists a model belonging to the same class that is a definable $\Gamma$-filtration of $\mathcal{M}$.
\end{defin}

\begin{lemma}
  $ $

\begin{enumerate}
  \item Let $\mathcal{L}$ be a complete normal modal logic. If $\operatorname{Frames}(\mathcal{L})$ admits filtration, then $\mathcal{L}$ has the finite model property.
  \item If the class of models $\operatorname{Mod}(\mathcal{L})$ admits filtration, then $\mathcal{L}$ has the finite model property and Kripke complete as well.
\end{enumerate}
\end{lemma}

\section{Filtration of Euclidean logics}

First of all, let us ensure that a filtration of an Euclidean frame is not necessary finite.
Let $[x] \sim_{\Gamma} [y]$ and $[x] \sim_{\Gamma} [z]$. Then for some $x' \in [x]$ $y' \in [y]$, one has
$x' R y'$ and $x'' R z'$ for some $x'' \in [x]$ and $z' \in [z]$. Clearly, we cannot claim that $x' = x''$ in general. Thus, minimal filtration does not preserve the required property.

\begin{lemma}
  ${\bf K5}$ admit filtration.
\end{lemma}

\begin{proof}
  Let $\mathcal{M}$ be a ${\bf K5}$-model and $\Gamma_0$ a finite set of formulas closed under subformulas.
  Let us put $\Gamma = \Gamma_0 \cup \operatorname{Sub}(\{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma_0 \}) \cup \Psi$, where $\Psi = \nabla_1 \nabla_2 \dots \nabla_n \Box \psi$ for $\Box \psi \in \Gamma_0$ and
  $\nabla_i \in \{ \Diamond, \Box \}$. By Proposition~\ref{K5proves}, any element of $\Phi$ has one of the four forms. Thus, $W \sim_{\equiv_{\Gamma}}$ has a finite index.
  We put $\widehat{R} = R^{\operatorname{max}}_{\Gamma}$.
\end{proof}

\begin{defin} A first-order formula is called Horn if it has the following form:

  \begin{center}
    $\forall x_1, \dots, x_n (x_{i_1} R x_{j_1} \land \dots \land x_{i_s} R x_{j_s} \rightarrow x_k R x_l)$
  \end{center}
\end{defin}

\begin{defin}
  Let $H$ be a Horn property and $\langle W, R \rangle$ a Kripke frame. A Horn closure of a binary relation $R$ is the minimal relation $R^{H}$ containing $R$ and satisfying $H$.
\end{defin}

\begin{lemma}
  $R^{H} = \bigcup \limits_{n < \omega} R_n$ where

  \begin{enumerate}
    \item $R_0 = R$.
    \item $R_{n + 1} = R_n \cup \{ (a, b) \in W \: | \: \exists \vec{c} \in W \: P(a, b, \vec{c})\}$, where $P$ is a premise of $H$.
  \end{enumerate}
\end{lemma}

$E$-closure (an Euclidean Horn closure of a binary relation) has the following equivalent definitions:
\begin{lemma} \label{equivHorn}
  Let $\mathcal{F} = \langle W, R \rangle$ be a Kripke frame.
  The following conditions are equivalent:

  \begin{enumerate}
    \item $R^{E}$ is the smallest Euclidean relation containing $R$.
    \item $R^{E} = \bigcup \limits_{i < \omega} R_i$, where
    \begin{itemize}
      \item $R_0 = R$
      \item $R_{n + 1} = R_n \cup (R^{-1}_n \circ R_n)$
    \end{itemize}
    \item $x R^E y$ iff there exists $n < \omega$ such that
    either $x R y$ or $\exists z_1, \dots, z_n$ with $z_1 R x$ and $z_{n - 1} R y$ and for each $1 < i \leq n$ one has
    either $z_{i - 1} R z_i$ or $z_i R z_{i - 1}$.
    \item $R^{E} = R \cup \bigcup \limits_{i < \omega} (R^{-1} \circ (R \circ R^{-1})^n \circ R)$.
  \end{enumerate}
\end{lemma}

\begin{proof}
$ $

\begin{enumerate}
  \item $(1) \Rightarrow (2)$
  Let us show that if $R^{E}$ is the smallest Euclidean relation containing $R$, then $R^{E} = \bigcup \limits_{i < \omega} R_i$. There are two inclusions:
  \begin{itemize}
    \item $R^{E} \subseteq \bigcup \limits_{i < \omega} R_i$. Recall that $R^{E}$ has the form (?):
    \begin{center}
      $R^{E} = \bigcap \{ R' \: | \: R \subseteq R', \forall a, b \in W \: R'(a,b) \Rightarrow \exists x \in W \: R'(x, a) \: \& \: R'(x, b)\}$
    \end{center}

    \item $\bigcup \limits_{i < \omega} R_i \subseteq R^{E}$.
    Let us show that $x R_n y$ for each $n < \omega$ implies $x R^E y$ by induction on $n$.
    If $n = 0$, then $x R y$, thus, $x R^{E} y$, since $R$ is a subrelation of $R^{E}$.
    Suppose $n = m + 1$ and $x R_{m + 1} y$. Let us show that $x R^{E} y$.
    From $x R_{m + 1} y$, one has $(x, y) \in R^n \cup (R^{-1}_n \circ R_n)$. There are two cases:
    \begin{itemize}
      \item $x R^n y$, one needs to merely apply the IH.
      \item $x R^{-1}_n \circ R_n y$. Then $\exists z \in W \: x R^{-1}_n z \: \& \: z R_n$. That is,
      $z R_n x$ and $z R_n y$ for some $z$. $R_n$ is already a subrelation of $R^E$. Thus,
      $z R^{E} x$ and $z R^{E} y$. That implies $x R^{E} y$.
    \end{itemize}
  \end{itemize}
  \item $(2) \Rightarrow (3)$
  Let $(x, y) \in R_m$, let us the statement by induction on $m$.
  \begin{enumerate}
    \item Suppose $m = 0$, then $x R y$, and the statement is shown putting $n = 0$.
    \item Suppose $m = p + 1$ and $x R_{p + 1} y$.
    Assume that either $x R y$ or $\exists z_1, \dots, z_p$ with $z_1 R x$ and $z_{p - 1} R y$ and for each $1 < i \leq p$ one has
    either $z_{i - 1} R z_i$ or $z_i R z_{i - 1}$.

    $x R_{p + 1} y$ implies $(x, y) \in R_{p} \cup (R^{-1}_p \circ R_p)$. If $(x, y) \in R_{p}$, then we merely apply the IH.
    Suppose $(x, y) \in R^{-1}_p \circ R_p$, then $(z, x) \in R_p$ and $(z, y) \in R_p$
  \end{enumerate}
  \item $(3) \Rightarrow (4)$
  Suppose either $x R y$ or there exist $n \geq 1$ and $z_1, \dots, z_n$ with $z_1 R x$ and $z_{n - 1} R y$ and for each $1 < i \leq n$ one has either $z_{i - 1} R z_i$ or $z_i R z_{i - 1}$.
  If $x R y$, then we are done. Otherwise there exists $n \geq 1$ with the condition above. Then $(x, y) \in R_{n + 1}$ that follows from the condition.
  \item $(4) \Rightarrow (1)$
\end{enumerate}
\end{proof}

\begin{lemma}
  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be an Euclidean model, $\Gamma$ a set of $\operatorname{Sub}$-closed formulas, and $\sim$ an equivalence relation having a finite index that respects $\Gamma$, then $\widehat{R} = (R^{min}_{\Phi})^{E} \subseteq R^{max}_{\Gamma}$, where $\Phi = \Gamma \cup \{ \Diamond \Box \varphi \: | \: \Box \varphi \in \Gamma \}$.

  Thus, ${\bf K}5$ admits strict filtrations.
\end{lemma}

\begin{proof}
  Recall that $(R^{min}_{\Phi})^{E}$ has the form $(R^{min}_{\Phi})^{E} = \bigcup \limits_{n < \omega} (R^{min}_{\Phi})_n$, where
  \begin{enumerate}
    \item $(R^{min}_{\Phi})_0 = R^{min}_{\Phi}$
    \item $(R^{min}_{\Phi})_{m + 1} = (R^{min}_{\Phi})_n \cup (((R^{min}_{\Phi})_n)^{-1} \circ (R^{min}_{\Phi})_n)$
  \end{enumerate}
  One needs to show that for each $n < \omega$ $(R^{min}_{\Phi})_n \subseteq R^{max}_{\Gamma}$. We prove this by induction.
  Suppose $\mathcal{M}, x \models \Box \varphi$ for $\Box \varphi \in \Phi$ and $[x] (R^{min}_{\Phi})^{E} [y]$. We need $\mathcal{M}, y \models \varphi$.
  \begin{enumerate}
    \item $([x], [y]) \in (R^{min}_{\Phi})_0$, then $([x], [y]) \in R^{min}_{\Phi}$. Then there exist $x' \in [x]$ and $y' \in [y]$ such that $x' R y'$. So $\mathcal{M}, x' \models \Box \varphi$ and, thus,
    $\mathcal{M}, y' \models \varphi$. Then $\mathcal{M}, y' \models \varphi$ as well since $y' \in [y]$.
    \item $([x], [y]) \in (R^{min}_{\Phi})_{m + 1}$, then
    $([x], [y]) \in (R^{min}_{\Phi})_{m} \cup (((R^{min}_{\Phi})_{m})^{-1} \circ R^{min}_{\Phi})_{m})$.

    If $([x], [y]) \in (R^{min}_{\Phi})_{m}$, then we apply the IH.

    Suppose $([x], [y]) \in (R^{min}_{\Phi})_{m})^{-1} \circ (R^{min}_{\Phi})_{m}$,
    then there exists $[z] \in W / \sim_{\Phi}$ such that $([z], [x]) \in (R^{min}_{\Phi})_{m})$ and $([z], [y]) \in (R^{min}_{\Phi})_{m}$.

    Then one has the following picture (using Lemma~\ref{equivHorn}):

    \xymatrix{
&&    [z] & \ar[l]_{R^{min}_{\Phi}} [z_1] \ar@{-}[r]^{R'} & [z_2] \ar@{-}[r]^{R'} & \ar@{-}[r]^{R'} \dots & [z_{m - 1}] \ar@{-}[r]^{R'} & [z_m] \ar[r]^{R^{min}_{\Phi}} & [x] \\
&&    [z] & \ar[l]^{R^{min}_{\Phi}} [z_1^{'}] \ar@{-}[r]_{R'} & [z_2^{'}] \ar@{-}[r]_{R'} & \ar@{-}[r]_{R'} \dots & [z_{m - 1}] \ar@{-}[r]_{R'} & [z_m^{'}] \ar[r]_{R^{min}_{\Phi}} & [y]
    }
  Where $R'$ is either $R^{min}_{\Phi}$ or its converse.
  One has $\mathcal{M}, x \models \Box \varphi$ for $\Box \varphi \in \Phi$, where $\widehat{M}$ is the minimal filtration of $\mathcal{M}$ through $\Phi$. One has $[z_m] R^{min}_{\Phi} [x]$, then $a_m R a$ for some $a_m \in [z_n]$ and $a \in [x]$. Thus, $\mathcal{M}, a_m \models \Diamond \Box \varphi$ and, thus, $\widehat{\mathcal{M}}, [z_m] \models \Diamond \Box \varphi$.

  Applying Proposition~\ref{k5useful} several times, one may show that $\widehat{\mathcal{M}}, [z_1] \models \Diamond \Box \varphi$.
  One has $[z_1] R^{min}_{\Phi} [z]$, then for some $a \in [z]$ and $a_1 \in [z_1]$ we have $a_1 R a$.

  Then $\mathcal{M}, a \models \Box \varphi$ and $\widehat{M}, [z] \models \Box \varphi$.

  We have $[z_1^{'}] R^{min}_{\Phi} [z]$, thus, $a_1^{'} R a^{'}$ for some $a_1^{'} \in [z_1^{'}]$ and $a^{'} \in [z]$. Then
  $\mathcal{M}, a_1^{'} \models \Diamond \Box \varphi$, and, thus, $\widehat{M}, [z_1^{'}] \models \Diamond \Box \varphi$.

  One may show that $\widehat{M}, [z_{m}^{'}] \models \Diamond \Box \varphi$ in the same way via Lemma~\ref{k5useful}. Thus, $\mathcal{M}, z_{m}^{'} \models \Diamond \Box \varphi$. We also have $\mathcal{M}, z_{m}^{'} \models \Diamond \Box \varphi \to \Box \varphi$, and, thus,
  $\mathcal{M}, z_{m}^{'} \models \Box \varphi$. Then $\widehat{\mathcal{M}}, [z_m^{'}] \models \Box \varphi$.

  One has $[z_m^{'}] R^{min}_{\Phi} [y]$, then $a_m^{'} R y'$ for some $a_m^{'} \in [z_m^{'}]$ and $y' \in [y]$. Then $\mathcal{M}, y' \models \varphi$. But $y' \sim_{\Phi} y$, so $\mathcal{M}, y \models \varphi$.
  \end{enumerate}
\end{proof}


\subsection{Clusters}

Let $\mathcal{F} = \langle W, R \rangle$ be a transitive frame.
Let us put $x R^{\bullet} y \Leftrightarrow x R y \: \& \: \neg (x R y)$. A point $x$ is proper if $x R x$. Let us define the following
equivalence relation:

\begin{center}
  $x \equiv y \Leftrightarrow x R y \: \& \: y R x \lor x = y$.
\end{center}
A cluster is an element of the quotient set $W / \equiv$. Given $x \in W$, $C_x$ is a cluster containing $x$. Thus $C_x =
\{ x\} \cup \{ y \: | x R y x \}$. The original relation lifts to the antisymmetric transitive relation on $W / \equiv$ defined as
$C_x R C_y$ iff $x R y$. A cluster $C$ is called maximal if $C R C'$ implies $C = C'$. A point is $R$-maximal if $C_x$ is a maximal cluster,
that is, $R^{\bullet}(x) = \emptyset$. A degenerated cluster is a singleton $\{ x \}$ with $\neg (x R x)$. A cluster is called simple if it has the form $\{ x \}$ with $x R x$. If $\langle W', R' \rangle$ is an inner substructure of $\langle W, R \rangle$, then every $R'$-cluster is an $R$-cluster and every $R$-cluster that intersects $W'$ is a subset of $W'$ and is an $R'$-cluster itself. Given a Kripke model $\mathcal{M}$, a set of formulas $\Gamma$ is satisfied by a cluster $C$ if every member of $\Gamma$ is true at some point of $C$.

If clusters coincide then their poitns have the same theory in the original model:

\begin{lemma}
  $C_x = C_y$ implies $\mathcal{M}, x \models \Box \varphi \Leftrightarrow \mathcal{M}, y \models \Box \varphi$
\end{lemma}

Let us describe the bulldozing technique allowing one to eliminate nondegenerated clusters \cite{bull1984basic}. Let $\mathcal{L}$ be a transitive logic and $\mathcal{F}$ its frame. We construct first a frame $\mathcal{F}^0 = \langle W^{0}, R^{0} \rangle$ replacing every nondegenerated frame $C$ of $W$ by $C^0 = \{ \langle x, i \rangle \: | \: x \in C, i < \omega \}$. We also replace each degenerated cluster $C$ by $\{ \langle x,0 \rangle\}$.
Elements of these subsets form $W^0$. The relation $R^{0}$ is defined as
\begin{center}
  $\langle x, i \rangle R^{0} \langle y, j \rangle \Leftrightarrow x R^{\bullet} y \text{ or } (x \equiv y \: \& \: i < j) \text{ or } i = j \: \& \: x <_C y$
\end{center}
where $<_C$ is an arbitrary strict ordering on the proper cluster $C$ containing $x$ and $y$.

Each nondegenrated cluster $C$ is replaced by an infinite set $C_0$ such that $\langle C_0, R_0 \rangle$ is a strict linear order. Moreover, $\langle y, j \rangle$, a copy of $y$, occurs after $\langle y, j \rangle$, a copy of $x$.

Bulldozing might be extended to models $\mathcal{M} = \langle W, R, \vartheta \rangle$ defining $\vartheta^0$ as follows
\begin{center}
  $\vartheta^{0}(p_i) = \{ \langle x, i \rangle \: | \: x \in \vartheta(p_i), i < \omega \}$.
\end{center}

One may show inductively the following fact.
\begin{lemma}
  $\mathcal{M}, x \models \varphi \Leftrightarrow \mathcal{M}^{0}, \langle x, i \rangle \models \varphi$
\end{lemma}

Let us concretise the case of transitive Euclidean frames. First of all, we consider clusters in ${\bf K}45$ frames.

\subsection{}

\section{Transitive closure stuff}

\section{PDLisation of Euclidean logics}

\section{Transitive closure and fusion}

\bibliographystyle{plain}
\bibliography{Text}

\end{document}
